---
title: "Analysis of 2025 Company Pledge Counterfactual Value Survey Results"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}
# Load relevant packages
pacman::p_load(tidyverse, dbplyr, gt, PHEindicatormethods,jsonlite)
```

```{r params}
theme_set(theme_bw())
theme_update(legend.position = "bottom", plot.title.position = "plot")

```

```{r import-processed-data}
# Import processed results
source("connect-parfit.R")

results <- tbl(con, in_schema("impact_surveys", "cpcv2025_clean")) %>% 
  collect() %>% 
  # Get number of years of pledger at time of survey completion
  mutate(
    start_date = mdy(start_date),
    years_of_pledge = interval(start_date, time_started) / years(1),
    gwwc_caused_usd = donations_total_usd-counterfactual_donations_total_usd)
```

```{r import-sample}
# Import sample

sample <- tbl(con, in_schema("impact_surveys", "cpcv2025_sample"))  %>% 
  collect() %>% 
  mutate(responded = id %in% results$id,
         completed = id %in% results$id[results$survey_completed])
```

## Document purpose

Here I will conduct quantitative analysis of the results of the 2025 Company Pledge Counterfactual Value Survey, which was run as part of Giving What We Can's internal impact evaluation for the 2023--2024 period. The main purpose of this document was to provide preliminary data on the impact of the Company Pledge.

## Survey response rates and representativeness

### Response rates

This survey sample was all companies with an active company pledge, whose company pledge was at least 1 year old at the sampling date.

However, after receiving our survey, several organisations responded letting us know that they wished to depledge or that they had ceased operations.


```{r}
get_response_rates <- function(x, by_col) {
  x %>% 
    summarise(
    sampled = n(),
    responded = sum(responded),
    completed = sum(completed),
     .by = {{by_col}}) %>% 
  phe_proportion(responded, sampled, "standard") %>% 
  rename_with(~str_c("response_",.x),.cols = c(value,uppercl,lowercl)) %>% 
  phe_proportion(completed, sampled, "standard") %>% 
  rename_with(~str_c("complete_",.x),.cols = c(value,uppercl,lowercl)) %>% 
  arrange({{by_col}}) 
}

make_response_rates_table <- function(x) {
  x %>% 
  gt() %>% 
  fmt_percent(
    c(starts_with("response_"), starts_with("complete_")),
    decimals = 0) %>% 
  cols_merge(starts_with("response_"),pattern = "{1}&nbsp;({2}&#8288;–&#8288;{3})")%>% 
  cols_merge(starts_with("complete_"),pattern = "{1}&nbsp;({2}&#8288;–&#8288;{3})") %>% 
  cols_label(
    sampled ~ "Number sampled",
    responded ~ "Number responded",
    completed ~ "Number completed",
    response_value ~ html("Response&nbsp;rate"),
    complete_value ~ html("Completion&nbsp;rate")
  )
}


response_rates <- sample %>%
  get_response_rates() 

response_rates  %>% 
  make_response_rates_table()
  
```

Response rates for this survey were around 30%, similar to that for our 2023 Pledger Reporting and Counterfactuality survey.

Among non-respondents, were a number of Company Pledgers, who informed us that they wished to depledge after they received the survey:

```{r}
label_columns <- \(x) str_to_sentence(str_replace_all(x,"_", " "))

sample %>% 
  summarise(
    num_sampled = n(),
    num_depledged = sum(!is.na(depledge_date)),
    num_responded = sum(responded),
    num_responded_or_depledged = sum(!is.na(depledge_date) | responded),
    percent_sample_responded_or_depledged = num_responded_or_depledged/ num_sampled,
    percent_sample_depledged = num_depledged / num_sampled,
    percent_respondents_depledged = num_depledged / num_responded_or_depledged
    ) %>% 
  gt(
    caption = "Summary of depledging among survey respondents"
  ) %>% 
  cols_label_with(fn = label_columns) %>% 
  fmt_percent(starts_with("percent"))
  
```

If we include those who depledged among respondents, then the response rate was closer to 40% and roughly 16% of respondents depledged. For the rest of this analysis we will only consider those who opened the survey to be respondents.


## Results

### Donations

First, we will consider the response to the first question, which asked respondents to estimate their total donations for the duration of their Company Pledge:

```{r}
results %>% 
  summarise(
    respondents = n(),
    mean_pledge_donations = mean(donations_total),
    median_pledge_donations = median(donations_total),
    total_pledge_donations = sum(donations_total),
  ) %>% 
  rename_with(label_columns) %>% 
  pivot_longer(cols = everything(), names_to = "Variable", values_to = "Value (USD)") %>% 
  gt(caption = "Summary statistics for question: Since your Company Pledge commenced on {start_date}, roughly *how much has your company donated to high-impact charities/funds as part of the pledge*?") %>% 
  fmt_currency(
    2,2:4, suffixing = T
  )
```

This is the only question that all respondents provided an answer to. Most respondents provided an answer to the remaining questions, but the respondent who reported the largest pledge donations (~$3M, ~40% of the total) did not. This means this respondents donations are missing from our evaluation, which likely has a significant effect on the rest of our quantitative analysis.

### Pledge adherence

This section will look into pledge adherence among respondents to the survey. This means just considering respondents who reported their profit.

First, let's get a breakdown of the results among those who did report a profit:

```{r}

results %>%
  filter(!is.na(total_profit_usd)) %>% 
  summarise(
    respondents = n(),
    total_pledge_years = sum(years_of_pledge),
    reported_any_profit = sum(total_profit_usd>0),
    combined_profit_usd = sum(total_profit_usd),
    mean_profit_USD = mean(total_profit_usd),
    average_profit_per_pledge_year = sum(total_profit_usd)/total_pledge_years
  ) %>% 
  rename_with(label_columns) %>% 
  pivot_longer(cols = everything(), names_to = "Variable", values_to = "Value") %>% 
  gt(
    caption = "Summary of profits among respondents to Company Pledge Value Survey",
    rowname_col = "Variable"
    ) %>% 
  fmt_currency(2, 4:6, suffixing = T) %>% 
  fmt_number(2,1:3,decimals = 1)  


```

Notably 10 of the 15 Companies reported making more than $0 in profit since their pledge started, leaving 33% of Company Pledgers without profits.

Now, considering the fraction who did report a profit, how much have they donated as a proportion of profit?

```{r}
results %>%
  filter(!is.na(total_profit_usd) & total_profit_usd >0) %>% 
  mutate(percent_donated = donations_total_usd/total_profit_usd) %>% 
  count(percent_donated) %>% 
  arrange(percent_donated) %>% 
  gt(caption = "Percent of profits donated (Company Pledge Respondents who reported a profit)") %>% 
  fmt_percent(1)
```

All but two respondents reported total donations equivalent to 10% of their profits or higher. The remaining two have donations between 5% and 10%.

### Counterfactual donations per company pledge year

Next, I will try and get some information on GWWC's counterfactual influence on the amount donated by Company Pledgers.

```{r}


results %>% 
  filter(!is.na(counterfactual_donations_total_usd)) %>% 
  summarise(
    num_pledgers = n(),
    across(c(donations_total_usd,gwwc_caused_usd, years_of_pledge),sum),
    counterfactual_donations_per_pledge_year = gwwc_caused_usd / years_of_pledge
    ) %>% 
  mutate(
    `Number of pledgers` = num_pledgers,
    `Total pledge years` = years_of_pledge,
    
    `Total pledge donations` = donations_total_usd,
    `Total GWWC-attributable donations` = gwwc_caused_usd,
    `Percent GWWC-attributable donations` = gwwc_caused_usd/ donations_total_usd,
    `Pledge donations per pledge year` = donations_total_usd / years_of_pledge,
    `GWWC-attributable donations per pledge year` = counterfactual_donations_per_pledge_year,
    .keep = "none") %>% 
  pivot_longer(cols = everything(), names_to = "Variable", values_to = "Value") %>% 
  gt(
    caption = "Summary of GWWC-attributable company pledge donations",
    rowname_col = "Variable"
    ) %>% 
  fmt_currency(2, c(3:4,6:7), suffixing = T) %>% 
  fmt_percent(2,5) %>% 
  fmt_number(2,1:2,decimals = 1)
  

```


### Other counterfactuality results

#### Counterfactual activities

First, what GWWC activities did respondents report influenced the amount they donated to high-impact charities and funds in 2024:

```{r}
default_options <- c("The Company Pledge, 🔸10% Pledge or 🔹Trial Pledge", "Our giving recommendations","The donation platform allowing you to donate to a wide range of charities","The donation platform allowing you to donate tax-deductibly.","Our research content (e.g., our cause pages, charity pages, why we recommend funds, etc.)","Our events (e.g. Effective Giving Day)","Members of the Giving What We Can community","Members of our team", "Other")

# Get results for those who responded to this question
responded_to_activities <- results %>% 
   filter(!is.na(counterfactual_activities))

responded_to_activities %>% 
  # reframe from JSON to list all activities
  reframe(activities = lapply(counterfactual_activities, function(s) {
    if (is.na(s)) return(NA)
    fromJSON(s)
    }) %>% unlist
) %>% 
  mutate(activities = factor(activities, default_options)) %>% 
  summarise(
    Respondents.Number = n(),
    Respondents.Percent = n()/nrow(responded_to_activities),
    .by = activities) %>% 
  arrange(activities) %>% 
  gt(
    caption = sprintf("Question: What part(s) of Giving What We Can have affected *how much* you donate to high-impact charities/funds? (Respondents: n = %i)",nrow(responded_to_activities)),
    rowname_col = "activities"
    ) %>% 
  tab_spanner_delim(".") %>% 
  fmt_percent(ends_with("Percent"))
```

#### Affected charity choice

```{r}

results %>% 
  filter(!is.na(affected_charity_choice)) %>% 
  count(affected_charity_choice) %>% 
  mutate(percent = n/sum(n)) %>% 
  gt(caption = "Question: Has Giving What We Can or its donation platform affected which charities or funds your company donates to?") %>% 
  fmt_percent(3) %>% 
  cols_label_with(fn = label_columns)

```


### Other questions

#### Dashboard interest


```{r}
results %>% 
  filter(!is.na(dashboard_interest)) %>% 
  count(dashboard_interest) %>% 
  mutate(percent = n / sum(n),
         dashboard_interest = factor(
           dashboard_interest, 
           c("Very likely", "Likely", "Neutral",
             "Unlikely", "Very unlikely", "Unsure"
             )
           )
         ) %>% 
  arrange(dashboard_interest) %>% 
  gt(caption = "Question: How likely would you be to use a dashboard that allowed your company to record its donations and track progress against your Company Pledge?") %>% 
  fmt_percent(3) %>% 
  cols_label_with(fn = label_columns)

```

#### Pledge benefits

```{r}
results %>% 
  filter(!is.na(pledge_benefits)) %>% 
  count(pledge_benefits) %>% 
  mutate(percent = n / sum(n),
         pledge_benefits = factor(
           pledge_benefits, 
           c("Yes", "No",  "Unsure")
           )
         ) %>% 
  arrange(pledge_benefits) %>% 
  gt(caption = "Question: Are there examples where you believe the Company Pledge was beneficial for your company (e.g., in your marketing)?") %>% 
  fmt_percent(3) %>% 
  cols_label_with(fn = label_columns)
```

