---
title: "Clean 2023 Major Pledge Counterfactual Value Survey Results"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}
# Load relevant packages
pacman::p_load(tidyverse, dbplyr, gt, glue)
```


```{r import-raw-data}
source("connect-parfit.R")

pledge_ids <- tbl(con, in_schema("pledges", "verified_active_pledge")) %>% 
  select(pledge_id = id, person_id)


raw_data <- tbl(con, in_schema("impact_surveys", "mpcv2025_raw")) %>% 
  # Join pledge ids
  left_join(pledge_ids, by = "person_id") %>% 
  collect()

```

## Filter out incomplete responses

First filter out survey runs, that do not constitute meaningful responses:

1) That can't be linked to a respondent (because they contain no pledge_id)
2) That don't have responses to the first question (and therefore no data)


```{r essential-filters}
# Get actual survey runs
actual_runs <- raw_data %>% 
  # Remove responses with no id or no response to the first question
  filter(!is.na(person_id) & !is.na(amount_effect))

```

## Filter to include just one run from each respondent

First run an automated process to figure out which runs should be kept where possible.

```{r get-run-info}

run_info <- actual_runs %>% 
  mutate(
    survey_completed = !is.na(completed),
    survey_started = !is.na(survey_start),
    .keep = "unused"
  ) %>% 
  # Create a decision variable to decide what to do with a run
  mutate(
    decision = case_when(
      n() == 1 ~ "keep", # When only one run for the person keep this run
      n_distinct(amount_effect) > 1 ~ "unclear", # When the person reports contradictory platform affects handle manually
      sum(survey_started) > 1 ~ "unclear", # When the person started the survey more than once handle manually
      survey_started ~ "keep", # Otherwise keep any runs where the survey was started
      any(survey_started) ~ "discard", # If another run for the person was started, but this one wasn't then discard
      row_number() == 1 ~ "keep", # If no runs were started, just keep the first run
      .default = "discard" # Discard all the rest
    ),
    .by = person_id
  )

run_info %>% 
  summarise(
    `Total survey runs` = n(),
    `Total completed` = sum(survey_completed),
    `Total respondents` = n_distinct(person_id),
    `Keep` = sum(decision == "keep"),
     `Discard` = sum(decision == "discard"),
    `Unclear` = sum(decision == "unclear")
    
  ) %>% 
  gt(caption = "Automated handling of survey runs") %>% 
  tab_spanner("Run decision", 4:6)

```

Because the number of respondents is not equal to the number of runs flagged to keep and there are two runs tagged 'unclear', we need to manually select a run.

Both runs started, but only one complete: they seem to have exactly the same responses. I will keep the completed run. Run to keep: 15237009


```{r get-runs}
cleaned_runs <- run_info %>% 
  filter(decision == "keep" | run == 15237009) %>% 
  select(-decision)
```

We now have `r nrow(cleaned_runs)` runs from `r n_distinct(cleaned_runs$person_id)` respondents, as we would expect.

## Correct erroneous results

One respondent reported that over the course of their pledge they had donated $100M. Because this was such a substantial amount and differed by more than an order of magnitude compared to the donations we had recorded, we followed up with this donor. They reported that the amount entered was likely a mistake and they had really given around 15 M to 20M.  We will update this run with the middle of these two estimates (17.5 M) (The amount is already in USD).

```{r}
# Correct run with incorrect info
corrected_runs <- cleaned_runs %>% 
  mutate(
    # Update erroneous run
    donations_total = case_when(
      run == 15337628 ~ 17.5e6, 
      .default = donations_total
      )
  )
```


## Add new variables

Next, I want to add some new variables to the data. Specifically, I want to add the age of the pledge at the time the survey was completed.

The variables we are adding/updating are:

- start_date: converting to date type
- pledge_age: a new variable that gives age of pledge in years at the time of the survey
- reported_total_usd, donations_total_usd, counterfactual_donations_usd: New variables that are USD-coverted versions of the base columns. 

Exchange rates from mid-2023 are used, based on the fact that our sample is partially composed of our largest recording donors for 2023. This is a fairly arbitrary choice, but unlikely to have a large effect on the outcome.

```{r}

# Get exchange rates
exchange_rates <- tbl(con, in_schema("public", "exchange_rate")) %>% 
  filter(date == as.Date("2023-07-01")) %>% 
  collect()

variable_added <- corrected_runs %>% 
  # Join mid_period exchange rates
  left_join(exchange_rates, by = "currency_code") %>% 
  mutate(
    # Convert monetary vars to USD
    across(c(reported_total, donations_total,counterfactual_donations_total),
           ~.x  / rate, .names = "{.col}_usd"),
    # Convert start date to date object
    start_date = mdy(start_date), 
    # Get pledge age in years
    pledge_age = interval(start_date, time_started) / years(1)
    )
```


## Tag anomalous runs

Finally, append some notes to runs that we want to flag in our database.

```{r note-runs}
note_runs <- tribble(
~run, ~note,
15337628, "This run has had donations_total updated to reflect a correction",
99991, "This response was added manually as the respondent reported their answer by email rather than by following the survey link. This respondent reported GWWC had no effect on the amount they donated to high-impact charities."
)

noted_runs <- variable_added %>% 
  left_join(note_runs, by = "run")
```


## Export

The completely processed results are then written to Parfit

```{r export-results}

source("write_parfit.R")

table_name <- "mpcv2025_clean"

dbWriteTable(write_con, name = Id(schema = "impact_surveys", table = table_name), 
             value = noted_runs, 
             row.names = FALSE, 
             overwrite = TRUE,
             append = FALSE)


time_now <- Sys.time() %>% as.character()

dbExecute(write_con, glue_sql(
  "COMMENT ON TABLE impact_surveys.{`table_name`} IS 'This table contains the survey responses collected in 2025 as part of the 2025 Major Pledge Counterfactual Value Survey. Responses have been filtered to include one response per respondent.
                        This version was created at: {`time_now`}
                        More information about this survey can be found in the protocol linked in the raw survey table';",
  .con = write_con
))
```

