---
title: "Non-response bias"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




```{r load-packaages}
pacman::p_load(binom, tidyverse, gt)

```

The purposes of this document is to run some very quick statistical tests comparing the results of the non-response followup survey to those of the original Pledge 2023 Recording Accuracy survey.

First, did respondents to the non-response followup survey differ significantly from respondents to the original survey in their likelihood of having accurate pledge donations? Lets run a fisher test:

```{r fisher-test-results}
# Create the table
survey_data <- tribble(
 ~accurate_records, ~inaccurate_records,
  20, 28, # Original survey
  6 , 0   # Non-response followup
)

# Run Fisher's exact test
fisher.test(survey_data)
```
Next, what are the confidence intervals on the proportion of the sample whose recorded donations were accurate? 

```{r }
bind_rows(
  list(
  # Group 1: 20 out of 48
  Original = binom.confint(x = 20, n = 48, methods = "exact", conf.level = 0.95),
  # Group 2: 6 out of 6
  `Non-response` = binom.confint(x = 6, n = 6, methods = "exact", conf.level = 0.95)
  ), .id = "Sample"
) %>%
  select(-method) %>% 
  gt() %>% 
  fmt_percent(4:6)
```
