---
title: "Clean 2023 Pledger Reporting Accuracy and Counterfactuality Survey Results"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}
# Load relevant packages
pacman::p_load(tidyverse, dbplyr, gt, glue)
```


```{r import-raw-data}
source("connect-parfit.R")

normalize_usd_sql <- function(amount_col, currency_col, date_str) {
  glue::glue_sql("normalize_amount_to_usd({`amount_col`}::numeric, {`currency_col`}, {date_str}::date)", .con = con)
}


raw_data <- tbl(con, in_schema("impact_surveys", "prac2023_raw")) %>% 
  # Add columns with USD conversion for monetary values
  mutate(
    reported_2023_usd = sql(normalize_usd_sql("reported_2023", "currency_code", '2023-07-01')),
    donations_2023_usd = sql(normalize_usd_sql("donations_2023", "currency_code", '2023-07-01')),
    counterfactual_donations_2023_usd = sql(normalize_usd_sql("counterfactual_donations_2023", "currency_code", '2023-07-01')),
    # I chose 2022-07-01 as the conversion date as this is roughly halfway between the average pledge_start_date of responses (2019-06-15) and when surveys were run
    reported_total_usd = sql(normalize_usd_sql("reported_total", "currency_code", '2022-07-01')),
    donations_total_usd = sql(normalize_usd_sql("donations_total", "currency_code", '2022-07-01'))
    ) %>% 
  collect() %>% 
  # If reported_2023_usd/reported_total_usd are NA, currency_code is NA and reported_2023/reported_total_usd is zero
  mutate(
    reported_2023_usd = if_else(reported_2023 == 0, 0, reported_2023_usd),
    donations_2023_usd = if_else(donations_2023 == 0, 0, donations_2023_usd),
    counterfactual_donations_2023_usd = if_else(counterfactual_donations_2023 == 0, 0, counterfactual_donations_2023_usd),
    reported_total_usd = if_else(reported_total == 0, 0, reported_total_usd),
    donations_total_usd = if_else(donations_total == 0, 0, donations_total_usd),
    # Round all USD values to nearest cent
    across(ends_with("_usd"), \(x) round(x, 2))
  ) 
  

```

## Filter out incomplete responses

First filter out survey runs, that do not constitute meaningful responses:

1) That can't be linked to a respondent (because they contain no pledge_id)
2) That don't have responses to the email question


```{r essential-filters}
# Get actual syrvey runs
actual_runs <- raw_data %>% 
  # Remove responses with no pledge id
  filter(!is.na(pledge_id) & !is.na(accurate_2023))

```

## Filter to include just one run from each respondent

First run an automated process to figure out which runs should be kept where possible.

```{r get-run-info}

run_info <- actual_runs %>% 
  mutate(
    survey_complete = !is.na(time_finised),
    survey_started = 
      # A started survey is one with the first question answered
      # First question differs depending on response to the email question
      !is.na(donations_2023) & accurate_2023 !="Yes" | 
      !is.na(counterfactual_donations_2023)
  ) %>% 
  # Create a decision variable to decide what to do with a run
  mutate(
    decision = case_when(
      n() == 1 ~ "keep", # When only one run for the person keep this run
      n_distinct(accurate_2023) > 1 ~ "unclear", # When the person reports contradictory platform affects handle manually
      sum(survey_started) > 1 ~ "unclear", # When the person started the survey more than once handle manually
      survey_started ~ "keep", # Otherwise keep any runs where the survey was started
      any(survey_started) ~ "discard", # If another run for the person was started, but this one wasn't then discard
      row_number() == 1 ~ "keep", # If no runs were started, just keep the first run
      .default = "discard" # Discard all the rest
    ),
    .by = pledge_id
  )

run_info %>% 
  summarise(
    `Total survey runs` = n(),
    `Total respondents` = n_distinct(pledge_id),
    `Keep` = sum(decision == "keep"),
     `Discard` = sum(decision == "discard"),
    `Unclear` = sum(decision == "unclear")
    
  ) %>% 
  gt(caption = "Automated handling of survey runs") %>% 
  tab_spanner("Run decision", 3:5)

```

Because the number of respondents is equal to the number of runs to keep and no run decisions are unclear, we can simply filter out the discards and have our responses.

```{r get-runs}
cleaned_runs <- run_info %>% 
  filter(decision == "keep") 
```

## Tag likely erroneously runs

Finally, add some notes to runs that seem to have anomalous results.

```{r}
note_runs <- tribble(
  ~run, ~notes,
  15010251, "Reported they gave nothing in 2023, but would have given more if they had never encountered GWWC. However, later in a freetext question they report they have actually donated a large amount equivalent to the amount they would have donated without GWWC to a DAF, which they intend to distribute to highly effective opportunities, in 2023. Implying GWWC hasn't really caused them to donate less.",
  15007973, "Reported they gave nothing, but would have given more if they had never encountered GWWC, but then added that they only started the pledge in 2024 not prior to 2023",
  15031022, "Responded to email saying they actually gave more than reported, but their donations_2023 is smaller than their reported 2023."
)

noted_runs <- cleaned_runs %>% 
  left_join(note_runs, by = "run")
```

## Export

The completely processed results are then written to Parfit

```{r export-results}

source("write_parfit.R")

table_name <- "prac2023_clean"

dbWriteTable(write_con, name = Id(schema = "impact_surveys", table = table_name), 
             value = noted_runs, 
             row.names = FALSE, 
             overwrite = TRUE,
             append = FALSE)


time_now <- Sys.time() %>% as.character()

dbExecute(write_con, glue_sql(
  "COMMENT ON TABLE impact_surveys.{`table_name`} IS 'This table contains the survey responses collected in 2025 as part of the 2023 10% Pledge Reporting Accuracy and Counterfactuality Survey. Responses have been filtered to include one response per respondent and to USD-normalised monetary values.
                        This version was created at: {`time_now`}
                        More information about this survey can be found in the protocol linked in the raw survey table';",
  .con = write_con
))
```

